Background for transient faults in asynchronous circuits

### 2019-11-23 draft_final.tex
ran spellcheck on ch 2

D. E. Muller proposed essentially this class of speed independent circuits, which are sometimes called Muller circuits \cite{Muller_59}.  Since then a number of synthesis methods following this design has been produced.  Usually the synthesis methods begin with some higher-level specification such as signal transition graph or Petri nets and translate into a state graph.  A state graph describes all the possible sequences of changes in a finite number of signals.  (The term wires and signals are used interchangeably as the state graph is converted into a circuit.)  Then depending on the underlying implementation, such as generalized C-element and standard C-implementations, the appropriate inputs and logic functions that allow the circuit to function according to the state graph is synthesized.  The end result is a hazard-free and speed independent asynchronous circuit.  The construction we describe begins with a semi-modular circuit, possibly resulting from the previously described process, and transforms it into a circuit that can tolerate any single transient fault.

@@

The rest of this work is organized as follows.  Chapter two introduces the different types of common faults and their causes, as well as the common fault models that researchers use when investigating these faults.  Also included in chapter two are the asynchronous circuit models that are relevant to this work.  In addition, there is discussion about the existing body of work on fault-tolerant solutions for both types of circuits with an emphasis on asynchronous circuits.  Chapter three provides the background terminology and definitions for asynchronous circuits.  Chapter four shows why the existing fault-tolerant methods are insufficient and propose a new fault-tolerant method for asynchronous transient faults.  We prove the correctness of the method using a combination of automated formal verification and manual methods.  And finally chapter five discusses how to better build fault-tolerant asynchronous circuits and suggests some possible new research directions.
%Overview:  why is the problem important  \\  maybe I can extend the Introduction a bit more

\chapter{Background}
\section{Faults}
Faults can be split into permanent and non-permanent faults.  Non-permanent faults are further split into transient and intermittent faults \cite{jha_gupta_2003}. %copied 
A fault is defined as the physical difference between the good or correct system and the current system.  An error may be caused by a fault, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function.  Hidden faults are also possible where the fault does not cause an error \cite{jha_gupta_2003}. \\

Permanent faults refer to the presence of a fault that affects the functional behavior of a system(circuit) permanently.  Some examples that may cause permanent faults are: incorrect connections between integrated circuits, broken components or parts of components, incorrect silicon and metal connections (manufacturing problem), and functional design errors.  Most permanent faults occur during fabrication.  There are many existing literature that cover permanent faults and the different tests at production time that can uncover the presence and location of these faults. \cite{jha_gupta_2003} \cite{giz_book_2006}.  We ignore the possible manufacturing problems and assume that the circuit is free from those problems.  Instead we consider the permanent faults that develop during operation from radiation.  For example, high energy ions can impact the transistors and the circuit causing short-circuiting, rupture of the gate oxide insulation and other permanent faults.\\ %copied from SEE eetimes article

Non-permanent faults is defined as faults that affect the system's functional behavior for a finite but unknown period of time.  They can occur at random moments and are categorized into two types, intermittent and transient faults.  Intermittent faults are caused by non-environmental conditions such as loose connections, deteriorating or aging components, critical timing (hazards and race conditions), resistance and capacitance variations (which may lead to timing faults), physical irregularities, and noise.  Intermittent faults usually affects the system for a shorter time period than the application time of a test developed for permanent faults and thus may not be detected using such a test.  Intermittent faults can transition into permanent faults naturally but may take a long time from hours up to months.  \\

@@

In this work we will not be concerned with intermittent faults since it can be thought of as a transition between transient to permanent faults.  Instead the two main fault modes we will look into are transient and permanent faults.  In both radiation effects, also called single event effects (SEEs), figure prominently.  SEEs are an umbrella term that covers the aforementioned SETs and SEUs as well as the single event effects causing permanent faults.  There are two main sources of radiation, one is cosmic rays from space which interact with the atmosphere to produce high energy neutrons.  Another is that $\alpha$-particles can be caused by trace impurities of radioactive elements present in the packaging material of the chip and the silicon substrate.  Radiation from cosmic rays become more severe at higher elevations while the latent source in packaging is difficult to eliminate.  Thus initially a lot of the work in fault-tolerant design was done for aviation and space flight.

@@

\section{Asynchronous Circuit and Synthesis}
In the previous chapter we described how an asynchronous circuit works from a higher level perspective and in particular that it should work under varying delays.  In this section we will cover some definitions and properties of the circuit and also how we can synthesize the circuit from a higher level specification.  Both Speed Independent and Delay Insensitive technologies will be introduced.  There are differences but also many similarities between the two methods.  Even though this work is based on the Speed Independent model, there has been some previous work done on the Quasi-delay Insensitive model so I wanted to provide some context to make a direct comparison.

@@

Next we describe the set of {\em transitions} of a circuit where only one signal value changes more formally.  
We write $s(w)$ for the value of signal $w$ in state $s$, and write $T(s, s')$ when there is a transition from $s$ to $s'$.

@@

%ADD SPECIFIC CONDITIONS THAT ALLOW SYNTHESIS
When the state graph have the following properties, it is possible to synthesize a hazard free and speed independent asynchronous circuit following the state graph.  Each variable from the state graph is implemented as a gate logic that takes other variables/wires as inputs and outputs a variable/wire.  We next look at how such a gate logic can be made.\\ 
\end{comment}

%C-elements
The C-element proposed by Muller is an important building block for asynchronous circuits.  It acts as a memory element in that it receives two inputs.  When both inputs are low, the C-element output goes low, and when both inputs are high the C-element output goes high.  However when the inputs are mismatched then the output gets held at its previous value.  This can be achieved in a number of ways such as a gate implementation as well as a transistor level implementation with feedback using a weak transistor, other implementations use capacitance to achieve this weak feedback.
\includegraphics[width=\textwidth]{celement}
\captionof{figure}{ a. logic of the C-element. A, B are inputs, C is the output \\  b. diagram of a C-element at the gate level   c. transistor-level implementation}

@@

The two specific Muller circuit implementations that we consider are the generalized C-element and standard C-implementations.  We assume that the circuits we want to make fault-tolerant are synthesized according to these two implementations.    
Each wire in the circuit is the output wire of a gate logic.  The gate logic has a set of input wires and the output wire changes value according to the values of the inputs.  We look at two conditions, set, the condition for the output wire to transition to a high value; and reset, the condition for the output wire to transition to a low value.  The gate logic then implements the set and reset conditions. 
We skim the details in finding the set and reset conditions from the circuit specifications. For each condition set and reset, there is an ON-set, OFF-set and Don't Care set.  ON-set are all input values where the wire is excited and a transition will occur.  The OFF-set includes all input values where the wire cannot be excited to some value per the circuit specifications.  Input values that are neither in the ON-set or OFF-set are in the Don't Care set.  So for the set condition, the ON-set are all input values where the wire must transition to high.  The OFF-Set includes all input values where the wire is excited instead to low, as well as input values for which the output wire must remain low.  The input values of the Don't Care set assign the output wire to be either high or low.  It does not matter which value is assigned since it does not disturb the function of the circuit and only helps for circuit optimization. \\

@@

Even though the generalized C-element is presented as a C-element with set and reset as inputs, the transistor level implementation differs slightly.  We assume a transistor level implementation of the set and reset combined with the C-element so that there is only one delay for the entire logic block.  In addition this is not a normal C-element because set is used exclusively for the pull-up logic and reset is used for the pull-down logic.  In a normal C-element we allow for any four combinations of the values of the set and reset input wires.  However using the pull-up and pull-down logic setup, we do not allow for when both set and reset is `ON' as the output wire would simultaneously be connected to the high voltage and ground.  Such an output would have indeterminable voltage values and violate the digital assumptions.  Thus, the set and reset logic is designed so that they cannot overlap.  This corresponds to $\overline{set}\vee\overline{reset} $ is always TRUE.\\
\begin{center}
\includegraphics[width=.7\textwidth]{genC}
\end{center}
\captionof{figure}{transistor level implementation of a generalized C-element.  The + signs are the set logic inputs and - signs are the reset logic inputs.  For this particular element, set is in3 AND in4 while reset is in1 AND in2.}

@@

For the standard C-implementation, we allow for more variation in delay.  For each wire in the circuit with a corresponding gate logic, we look at the set and reset logic as composition of various gates (such as AND, OR, NOT).  Each of these gates can have their own delay that is arbitrary in length that is instantized as an intermediate wire.  Thus the logic must be synthesized more carefully to eliminate hazards that can occur in these intermediate wires.  In particular the set and reset must be activated at states only when the wire is appropriately excited.  As a result, the standard C-implementation has a number of set logic functions combined by an OR gate and fed into one input of the C-element.  There will also be a number of reset logic functions combined by an OR gate and passed through an inverter and fed into another input of the C-element.  This implementation uses the standard C-element so that if the set and reset inputs are mismatched, the output is held.  In general the synthesis can be changed so that there is a hazard-free implementation using only two-input gates.

 %explain sop?

 \subsection{Quasi Delay Insensitive Circuits}
 %QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits.  Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays.  In particular each branch of a fork will have different delays and can be treated as separate wires.  One result about Delay Insensitive circuits is that it is a very restricted class.  Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed.  Note that in the case where all forks are isochronic, then this becomes the same as speed independent circuits.  \\

The gates of a QDI circuit is formed from Production Rules (PRs).  For the single output case, a gate takes a set of variables as inputs and outputs a variable.  Again a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$.  Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
   B_u\mapsto x\uparrow \\
   B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards.  Each guard is a boolean condition formed by some combination of the input variables.  When a guard is evaluated to be $True$ then the transition is activated.  When it is $False$ then the corresponding transition does not happen.  Note that similar to the generalized C-elements, both guards cannot be active at the same time.  In addition it is assumed that each guard is in a sum of products of the input variables.  The PRs of a circuit are evaluated all at the same time.  At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$.  This is due to the assumption of the arbitrary delay of a gate.  Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allow any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

@@

%soft error protection <- error correcting codes for SRAM memory
Because faults cannot be eliminated in digital circuits, different methodologies have been used to make the circuits fault-tolerant.  Much of the research done in this area are for synchronous circuits.  There has also been some research for asynchronous circuits using some similar techniques. First we present an overview of some common techniques for synchronous circuits, then introduce the current techniques for asynchronous circuits.  Some of the techniques in asynchronous design borrows or are inspired by the much more developed fault-tolerant synchronous circuits field.  However due to the difference in timing assumptions of the two types of circuits the techniques are not completely transferable.  We describe why some of the solutions from synchronous circuits cannot be directly applied to asynchronous circuits.\\

%- used in the early days for space and flight, and systems that CANNOT go offline/errors like stocks and banking.
%- different techniques... explain a bit about each
%-hardware, software, hybrid, 
%-detection vs correction
%- ECC/ parity checking (needs software/ controller) (Mitra's method)
%-TMR is still super important (possibly why TMR is not viable for async)
In the early days there were certain systems for which eliminating faults and errors were very important.  These include space and aviation control systems where due to the altitude there are more radiation and thus more faults.  In addition faults that cause error is very detrimental to these systems and may cause failures leading to loss of life.  Other examples of systems requiring reliable circuits include banking and stock market applications where the system must function without error for hours at a time and in case of any error may have devastating consequences, such as a bit flip of the most significant bit in a money transaction. \\

Solutions for eliminating these transient fault induced errors include physical, hardware, software or a mix of the previous.  Some commonly used techniques are transistor sizing, circuit hardening, hardware duplication, and time redundancy \cite{Mitra_08_softerrors}.  Transistor sizing is a physics based technique and does not reduce as much soft errors as the other techniques.  Circuit hardening is on the physics level, where it uses intrinsic properties such as resistive or capacitive hardening, or combining the transistors differently to have less effects from transient faults. (?)  Hardware duplication, is a common technique that tries to eliminate soft errors on the circuit level.  This is achieved by having multiple duplicates of a circuit and often comparing the value from each duplicate to find and eliminate errors.  For example Triple Modular Redundancy (TMR) has been used since the beginning of the field and is still a good solution for eliminating single soft errors.  The advantages are that it is able to eliminate nearly all soft errors but also due to the duplication, the area for each circuit and additional power increases substantially (correlates to the number of additional duplicates).  Time redundancy techniques re-use the same circuitry to compute values multiple times to ensure there is no error.  This technique is not completely hardware based and requires software to manage re-using the same circuity with the same inputs over time.  The area of the circuit does not change but the speed of the circuit worsens substantially. \\

There is also the additional distinction between error detection versus error correction.  In general it is easier to detect an error than to correct it.  For example using hardware duplication, only 2 copies are required to detect an error but 3 copies (TMR) to correct the error.  For different parts of the system there are also different strategies.  For example in memory components such as SRAM, soft error correction is usually done using error-correction codes.  These codes are compared to the bits that are stored to ensure that there are no errors.  They are designed to handle a predefined maximum number of errors.  For example Hamming codes where each codeword are at least a distance of 3 apart can handle one error.  This method is not purely hardware based and a controller system and software is used to match the codes and detect and correct any errors.  For a control circuit, there are also a variety of strategies.  Again, one could make use of error correcting codes or parity bits to detect or correct errors in combinational logic \cite{McCluskey_99}.  This uses a combination of hardware and software.  A completely software approach is to run a program on some input, then run the program again on a transformed version of the input where the output can be mapped back to the original output \cite{Mitra_softwareerror}.  If the output and the transformed output do not match then the fault is detected.  Lastly, for hardware only methods, even now TMR is a commonly used method because of its robustness and ensures error correction. \\
%However these cannot be used for latches or flip flops.  In these circuits,    Some of these techniques only detect errors but do not correct them.\\

Triple Modular Redundancy is a technique where each combinational logic unit is copied three times and then fed into a latch that performs a majority voting operation.  That is, if at least 2 out of the 3 signals agree on a value, the output becomes that value.  Additionally, to counter possible faults on the single output wire, the majority vote unit can also be copied three times and it's output wire fed appropriately as input to the three copies of the downstream combinational logic.  This would resolve both transient and permanent single faults.  However the downside is that the circuit needs to be copied three times so the area is three times larger and the power consumption as well.  In addition there must be many adaptations for this to work in asynchronous circuits as the signals can arrive in different orders.  We give an example of why TMR does not work in an asynchronous circuit as it does in a synchronous circuit below.  We assume that in the original circuit that each logic block has some input wires and an output wire.  In the TMR scheme, each logic block is triplicated and to remove single points of failure, we also triplicate the majority voter that the three copies of the output wire feeds into.  There are also three distinct copies of each input signal that correspond to outputs from the three voters.  This is the basic scheme that synchronous circuits with TMR uses.  Because this circuit is asynchronous, we assume that each logic block is composed of gates that has some arbitrary delay that is independent of other gates.  This is true under both delay assumptions of speed independent and delay insensitive circuits.  We ignore whether a fault can be in the voter for now. 
\begin{center}
\includegraphics[width=\textwidth]{tmrcounterex}
\end{center}
\captionof{figure}{An example of an error if a fault occurs in an asynchronous TMR circuit}
Initially all output wire values are `0's.  Then since the input signals are distinct copies of each other, it is possible for one set to finish transitions (new) before the other two sets and the first logic block transitions to a `1'.  Then a transient fault arrives on the output of the second logic block which transitions to a `1' as well.  The majority voter would then transition to a `1'.  However the upset from transient faults usually happen within a short time period and the old inputs eventually force the output of the second logic block back to `0'.  Then the majority voter would transition to a `0' so that causes a hazard on the output of the majority voter.  %depends on if the logic gate was combinational logic.... which I guess we can assume it is
%NEEDS FIXING AND CHECKING
It is also possible for the `1' from the majority voter to feed into other logic units which is effectively an early transition.  That early transition will then put the circuit into states that are not specified by the original state graph and thus the next states cannot be predicted.
\\

@@

Verdel and Makris proposes that simply porting the duplication strategy for error detection from synchronous circuits does not work for asynchronous circuits \cite{async_dup_ced}.  This is because the duplication strategy relies on both the original and duplicate to output the same value.  Some modifications that they propose to make the duplication error detection strategy work is to assume additional timing requirements.  That both copies should output the same value within a certain time determined by a pre-set delay element.  Everytime one of the signals transitions to a new value, and the duplicates are mismatched the timer starts and if they stays mismatched when the timer ends an error is detected.  They look at stuck-at-faults and premature firing faults.  There are also cases when the timer detection does not work and for those cases they propose using additional hardware to look at the states of all the wires to see if an illegal state was reached.  They present an example of a control circuit for data computations.  The main restrictions of this method is the additional timing assumptions they make about two copies of a circuit, and that the extra hardware components are assumed to be fault-free.  Also it requires an additional comparison unit for each pair of signals which may be costly.  Only one copy of the duplicates produce output signals that interfaces with the other units so that if there is an error on that signal it is not detected until some time after the error has propagated.
\\

@@

\chapter{New Background}
\section{Assumptions:  Boolean signals}
We make the assumption that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the the gates and C-elements are excited, and their outputs transition, we assume that transition is instantaneous and that after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.

>>>>

D. E. Muller proposed essentially this class of speed independent circuits, which are sometimes called Muller circuits \cite{Muller_59}.  Since then a number of synthesis methods following this design has been produced.  Usually the synthesis methods begin with some higher-level specification such as a signal transition graph or Petri net and translate into a state graph.  A state graph describes all the possible sequences of changes in a finite number of signals.  (The term wires and signals are used interchangeably as the state graph is converted into a circuit.)  Then depending on the underlying implementation, such as generalized C-element and standard C-implementations, the appropriate inputs and logic functions that allow the circuit to function according to the state graph is synthesized.  The end result is a hazard-free and speed independent asynchronous circuit.  The construction we describe begins with a semi-modular circuit, possibly resulting from the previously described process, and transforms it into a circuit that can tolerate any single transient fault.

@@

The rest of this work is organized as follows.  Chapter two introduces the different types of common faults and their causes, as well as the common fault models that researchers use when investigating these faults.  Also included in chapter two are the asynchronous circuit models that are relevant to this work.  In addition, there is discussion about the existing body of work on fault-tolerant solutions for both types of circuits with an emphasis on asynchronous circuits.  Chapter three provides the background concepts and definitions for asynchronous circuits.  Chapter four shows why the existing fault-tolerant methods are insufficient and propose a new fault-tolerant method for asynchronous transient faults.  We prove the correctness of the method using a combination of automated formal verification and manual methods.  And finally chapter five discusses how to better build fault-tolerant asynchronous circuits and suggests some possible new research directions.
%Overview:  why is the problem important  \\  maybe I can extend the Introduction a bit more

\chapter{Related Works}
\section{Faults}
Faults can be split into permanent and non-permanent faults. Non-permanent faults are further split into transient and intermittent faults \cite{jha_gupta_2003}. %copied 
A fault is defined as the physical difference between the good or correct system and the current system. An error may be caused by a fault, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function. Hidden faults are also possible where the fault does not cause an error \cite{jha_gupta_2003}. \\

Permanent faults refer to the presence of a fault that affects the functional behavior of a system(circuit) permanently. Some examples that may cause permanent faults are: incorrect connections between integrated circuits, broken components or parts of components, incorrect silicon and metal connections (manufacturing problem), and functional design errors. Most permanent faults occur during fabrication. There are many existing literatures that cover permanent faults and the different tests at production time that can uncover the presence and location of these faults. \cite{jha_gupta_2003} \cite{giz_book_2006}. We ignore the possible manufacturing problems and assume that the circuit is free from those problems. Instead we consider the permanent faults that develop during operation from radiation. For example, high energy ions can impact the transistors and the circuit causing short-circuiting, rupture of the gate oxide insulation and other permanent faults.\\ %copied from SEE eetimes article

Non-permanent faults are defined as faults that affect the system's functional behavior for a finite but unknown period of time. They can occur at random moments and are categorized into two types, intermittent and transient faults. Intermittent faults are caused by non-environmental conditions such as loose connections, deteriorating or aging components, critical timing (hazards and race conditions), resistance and capacitance variations (which may lead to timing faults), physical irregularities, and noise. Intermittent faults usually affect the system for a shorter time period than the application time of a test developed for permanent faults and thus may not be detected using such a test. Intermittent faults can transition into permanent faults naturally but may take a long time from hours up to months. \\

@@

In this work we will not be concerned with intermittent faults since it can be thought of as a transition between transient to permanent faults. Instead the two main fault modes we will look into are transient and permanent faults. In both radiation effects, also called single event effects (SEEs), figure prominently. SEEs are an umbrella term that covers the aforementioned SETs and SEUs as well as the single event effects causing permanent faults. There are two main sources of radiation, one is cosmic rays from space which interact with the atmosphere to produce high energy neutrons. Another is that $\alpha$-particles can be caused by trace impurities of radioactive elements present in the packaging material of the chip and the silicon substrate. Radiation from cosmic rays become more severe at higher elevations while the latent source in packaging is difficult to eliminate. Thus, initially a lot of the work in fault-tolerant design was done for aviation and space flight.

@@

\section{Asynchronous Circuit and Synthesis}
In the previous chapter we described how an asynchronous circuit works from a higher-level perspective and in particular that it should work under varying delays. In this section we will cover some definitions and properties of the circuit and also how we can synthesize the circuit from a higher-level specification. Both Speed Independent and Delay Insensitive technologies will be introduced. There are differences but also many similarities between the two methods. Even though this work is based on the Speed Independent model, there has been some previous work done on the Quasi-delay Insensitive model so I wanted to provide some context to make a direct comparison.

@@

Next, we describe the set of {\em transitions} of a circuit where only one signal value changes more formally. 
We write $s(w)$ for the value of signal $w$ in state $s$, and write $T(s, s')$ when there is a transition from $s$ to $s'$.

@@

%ADD SPECIFIC CONDITIONS THAT ALLOW SYNTHESIS
When the state graph has the following properties, it is possible to synthesize a hazard free and speed independent asynchronous circuit following the state graph. Each variable from the state graph is implemented as a gate logic that takes other variables/wires as inputs and outputs a variable/wire. We next look at how such a gate logic can be made.\\ 
\end{comment}

%C-elements
The C-element proposed by Muller is an important building block for asynchronous circuits. It acts as a memory element in that it receives two inputs. When both inputs are low, the C-element output goes low, and when both inputs are high the C-element output goes high. However, when the inputs are mismatched then the output gets held at its previous value. This can be achieved in a number of ways such as a gate implementation as well as a transistor level implementation with feedback using a weak transistor, other implementations use capacitance to achieve this weak feedback.
\includegraphics[width=\textwidth]{celement}
\captionof{figure}{a. logic of the C-element. A, B are inputs, C is the output \\ b. diagram of a C-element at the gate level c. transistor-level implementation}

@@

The two specific Muller circuit implementations that we consider are the generalized C-element and standard C-implementations. We assume that the circuits we want to make fault-tolerant are synthesized according to these two implementations. 
Each wire in the circuit is the output wire of a gate logic. The gate logic has a set of input wires and the output wire changes value according to the values of the inputs. We look at two conditions, set, the condition for the output wire to transition to a high value; and reset, the condition for the output wire to transition to a low value. The gate logic then implements the set and reset conditions. 
We skim the details in finding the set and reset conditions from the circuit specifications. For each condition set and reset, there is an ON-set, OFF-set and Don't Care set. ON-set are all input values where the wire is excited and a transition will occur. The OFF-set includes all input values where the wire cannot be excited to some value per the circuit specifications. Input values that are neither in the ON-set or OFF-set are in the Don't Care set. So, for the set condition, the ON-set are all input values where the wire must transition to high. The OFF-Set includes all input values where the wire is excited instead to low, as well as input values for which the output wire must remain low. The input values of the Don't Care set assign the output wire to be either high or low. It does not matter which value is assigned since it does not disturb the function of the circuit and only helps for circuit optimization. \\

@@

Even though the generalized C-element is presented as a C-element with set and reset as inputs, the transistor level implementation differs slightly. We assume a transistor level implementation of the set and reset combined with the C-element so that there is only one delay for the entire logic block. In addition, this is not a normal C-element because set is used exclusively for the pull-up logic and reset is used for the pull-down logic. In a normal C-element we allow for any four combinations of the values of the set and reset input wires. However, using the pull-up and pull-down logic setup, we do not allow for when both set and reset is `ON' as the output wire would simultaneously be connected to the high voltage and ground. Such an output would have indeterminable voltage values and violate the digital assumptions. Thus, the set and reset logic is designed so that they cannot overlap. This corresponds to $\overline{set}\vee\overline{reset} $ is always TRUE.\\
\begin{center}
\includegraphics[width=.7\textwidth]{genC}
\end{center}
\captionof{figure}{transistor level implementation of a generalized C-element. The + signs are the set logic inputs and - signs are the reset logic inputs. For this particular element, set is in3 AND in4 while reset is in1 AND in2.}

@@

For the standard C-implementation, we allow for more variation in delay. For each wire in the circuit with a corresponding gate logic, we look at the set and reset logic as composition of various gates (such as AND, OR, NOT). Each of these gates can have their own delay that is arbitrary in length that is instantized as an intermediate wire. Thus, the logic must be synthesized more carefully to eliminate hazards that can occur in these intermediate wires. In particular the set and reset must be activated at states only when the wire is appropriately excited. As a result, the standard C-implementation has a number of set logic functions combined by an OR gate and fed into one input of the C-element. There will also be a number of reset logic functions combined by an OR gate and passed through an inverter and fed into another input of the C-element. This implementation uses the standard C-element so that if the set and reset inputs are mismatched, the output is held. In general, the synthesis can be changed so that there is a hazard-free implementation using only two-input gates.
%explain sop?
\subsection{Quasi Delay Insensitive Circuits}
%QDI
Quasi Delay Insensitive Circuits (QDI) are related to Delay Insensitive Circuits. Previously we mentioned that Delay Insensitive circuits operate under the assumption that both the gate and the wire have unbounded gate delays. In particular each branch of a fork will have different delays and can be treated as separate wires. One result about Delay Insensitive circuits is that it is a very restricted class. Instead if that condition is relaxed so that some of the forks are allowed to be isochronic, then a more general class of circuits can be constructed. Note that in the case where all forks are isochronic, then this becomes the same as speed independent circuits. \\
The gates of a QDI circuit is formed from Production Rules (PRs). For the single output case, a gate takes a set of variables as inputs and outputs a variable. Again, a transition on the variable $x$ is denoted as $x\uparrow$, a transition from $x=0$ to $x=1$; and $x\downarrow$, a transition from $x=1$ to $x=0$. Then the two PRs associated with the gate with output $x$ is :
\begin{align*}
B_u\mapsto x\uparrow \\
B_d\mapsto x\downarrow 
\end{align*}
Where $B_u$ and $B_d$ are called guards. Each guard is a boolean condition formed by some combination of the input variables. When a guard is evaluated to be $True$ then the transition is activated. When it is $False$ then the corresponding transition does not happen. Note that similar to the generalized C-elements, both guards cannot be active at the same time. In addition, it is assumed that each guard is in a sum of products of the input variables. The PRs of a circuit are evaluated all at the same time. At any point in time, if a transition is activated, it will transition in a finite amount of time as long as the corresponding guard is held at $True$. This is due to the assumption of the arbitrary delay of a gate. Note that again, like the generalized C-element implementation, we assume that the transistor level implementation of a PR allows any arbitrary sum of products to be made where there are no delays within the logic of the sum of products.\\

@@

%soft error protection <- error correcting codes for SRAM memory
Because faults cannot be eliminated in digital circuits, different methodologies have been used to make the circuits fault-tolerant. Much of the research done in this area are for synchronous circuits. There has also been some research for asynchronous circuits using some similar techniques. First, we present an overview of some common techniques for synchronous circuits, then introduce the current techniques for asynchronous circuits. Some of the techniques in asynchronous design borrows or are inspired by the much more developed fault-tolerant synchronous circuits field. However due to the difference in timing assumptions of the two types of circuits the techniques are not completely transferable. We describe why some of the solutions from synchronous circuits cannot be directly applied to asynchronous circuits.\\

%- used in the early days for space and flight, and systems that CANNOT go offline/errors like stocks and banking.
%- different techniques... explain a bit about each
%-hardware, software, hybrid, 
%-detection vs correction
%- ECC/ parity checking (needs software/ controller) (Mitra's method)
%-TMR is still super important (possibly why TMR is not viable for async)
In the early days there were certain systems for which eliminating faults and errors were very important. These include space and aviation control systems where due to the altitude there are more radiation and thus more faults. In addition, faults that cause error is very detrimental to these systems and may cause failures leading to loss of life. Other examples of systems requiring reliable circuits include banking and stock market applications where the system must function without error for hours at a time and in case of any error may have devastating consequences, such as a bit flip of the most significant bit in a money transaction. \\

Solutions for eliminating these transient fault-induced errors include physical, hardware, software or a mix of the previous. Some commonly used techniques are transistor sizing, circuit hardening, hardware duplication, and time redundancy \cite{Mitra_08_softerrors}. Transistor sizing is a physics-based technique and does not reduce as much soft errors as the other techniques. Circuit hardening is on the physics level, where it uses intrinsic properties such as resistive or capacitive hardening, or combining the transistors differently to have less effects from transient faults. (?) Hardware duplication, is a common technique that tries to eliminate soft errors on the circuit level. This is achieved by having multiple duplicates of a circuit and often comparing the value from each duplicate to find and eliminate errors. For example, Triple Modular Redundancy (TMR) has been used since the beginning of the field and is still a good solution for eliminating single soft errors. The advantages are that it is able to eliminate nearly all soft errors but also due to the duplication, the area for each circuit and additional power increases substantially (correlates to the number of additional duplicates). Time redundancy techniques re-use the same circuitry to compute values multiple times to ensure there is no error. This technique is not completely hardware based and requires software to manage re-using the same circuity with the same inputs over time. The area of the circuit does not change but the speed of the circuit worsens substantially. \\

There is also the additional distinction between error detection versus error correction. In general, it is easier to detect an error than to correct it. For example, using hardware duplication, only 2 copies are required to detect an error but 3 copies (TMR) to correct the error. For different parts of the system there are also different strategies. For example, in memory components such as SRAM, soft error correction is usually done using error-correction codes. These codes are compared to the bits that are stored to ensure that there are no errors. They are designed to handle a predefined maximum number of errors. For example, Hamming codes where each codeword are at least a distance of 3 apart can handle one error. This method is not purely hardware based and a controller system and software are used to match the codes and detect and correct any errors. For a control circuit, there are also a variety of strategies. Again, one could make use of error correcting codes or parity bits to detect or correct errors in combinational logic \cite{McCluskey_99}. This uses a combination of hardware and software. A completely software approach is to run a program on some input, then run the program again on a transformed version of the input where the output can be mapped back to the original output \cite{Mitra_softwareerror}. If the output and the transformed output do not match then the fault is detected. Lastly, for hardware only methods, even now TMR is a commonly used method because of its robustness and ensures error correction. \\
%However these cannot be used for latches or flip flops. In these circuits, some of these techniques only detect errors but do not correct them.\\

Triple Modular Redundancy is a technique where each combinational logic unit is copied three times and then fed into a latch that performs a majority voting operation. That is, if at least 2 out of the 3 signals agree on a value, the output becomes that value. Additionally, to counter possible faults on the single output wire, the majority vote unit can also be copied three times and its output wire fed appropriately as input to the three copies of the downstream combinational logic. This would resolve both transient and permanent single faults. However, the downside is that the circuit needs to be copied three times so the area is three times larger and the power consumption as well. In addition, there must be many adaptations for this to work in asynchronous circuits as the signals can arrive in different orders. We give an example of why TMR does not work in an asynchronous circuit as it does in a synchronous circuit below. We assume that in the original circuit that each logic block has some input wires and an output wire. In the TMR scheme, each logic block is triplicated and to remove single points of failure, we also triplicate the majority voter that the three copies of the output wire feeds into. There are also three distinct copies of each input signal that correspond to outputs from the three voters. This is the basic scheme that synchronous circuits with TMR uses. Because this circuit is asynchronous, we assume that each logic block is composed of gates that has some arbitrary delay that is independent of other gates. This is true under both delay assumptions of speed independent and delay insensitive circuits. We ignore whether a fault can be in the voter for now. 
\begin{center}
\includegraphics[width=\textwidth]{tmrcounterex}
\end{center}
\captionof{figure}{An example of an error if a fault occurs in an asynchronous TMR circuit}
Initially all output wire values are `0's. Then since the input signals are distinct copies of each other, it is possible for one set to finish transitions (new) before the other two sets and the first logic block transitions to a `1'. Then a transient fault arrives on the output of the second logic block which transitions to a `1' as well. The majority voter would then transition to a `1'. However, the upset from transient faults usually happen within a short time period and the old inputs eventually force the output of the second logic block back to `0'. Then the majority voter would transition to a `0' so that causes a hazard on the output of the majority voter. %depends on if the logic gate was combinational logic.... which I guess we can assume it is
%NEEDS FIXING AND CHECKING
It is also possible for the `1' from the majority voter to feed into other logic units which is effectively an early transition. That early transition will then put the circuit into states that are not specified by the original state graph and thus the next states cannot be predicted.
\\

@@

Verdel and Makris proposes that simply porting the duplication strategy for error detection from synchronous circuits does not work for asynchronous circuits \cite{async_dup_ced}. This is because the duplication strategy relies on both the original and duplicate to output the same value. Some modifications that they propose to make the duplication error detection strategy work is to assume additional timing requirements. That both copies should output the same value within a certain time determined by a pre-set delay element. Everytime one of the signals transitions to a new value, and the duplicates are mismatched the timer starts and if they stay mismatched when the timer ends an error is detected. They look at stuck-at-faults and premature firing faults. There are also cases when the timer detection does not work and for those cases, they propose using additional hardware to look at the states of all the wires to see if an illegal state was reached. They present an example of a control circuit for data computations. The main restrictions of this method are the additional timing assumptions they make about two copies of a circuit, and that the extra hardware components are assumed to be fault-free. Also, it requires an additional comparison unit for each pair of signals which may be costly. Only one copy of the duplicates produces output signals that interfaces with the other units so that if there is an error on that signal it is not detected until some time after the error has propagated.
\\

@@

\chapter{Asynchronous Circuit Concepts \& Definitions}
\section{Assumptions:  Boolean signals}
We make the assumption that all signals are Boolean, that is, the signals are either `logic 0' or logic 1'.  When the the gates and C-elements are excited, and their outputs transition, we assume that transition is instantaneous and that after the transition, the signals are still Boolean.   
Even in cases where the wires are affected by faults, we assume that the new value of the wires are Boolean, and if they are inputs into circuit elements, that the output of the circuit elements also remain Boolean.  For each circuit, there is a finite set of wires, which we call $W$.  An exact definition $W$ will be presented at a later section, but for now just think of $W$ as including all the wires of some circuit.
